---
title: "Assignment 1"
output: html_notebook
author: "Ankit Mithbavkar"

---

## The Data

For this research project, I am using the iris data set.

This data set includes data for 50 flowers from each of 3 species of iris: setosa, versicolor, and virginica. The measurements are in centimeters for  the variables sepal length, sepal width, petal length, and petal width.


```{r}
library(datasets)
head(iris)
```

## Research Questions

1) How are sepal length and petal length related?

2) Which species of iris has the greatest average petal length?

## Plots

### Question 1

```{r}
iris |>
  ggplot(mapping = aes(x = Sepal.Length, y = Petal.Length)) + geom_point()
```


### Question 2

```{r}
iris |>
  group_by(Species) |>
  summarize(avg_petal_length = mean(Petal.Length, na.rm = TRUE)) |>
  ggplot(mapping = aes(x = Species, y = avg_petal_length)) + geom_bar(stat = "identity")
```

## Conclusion

### Question 1

Based off the scatter plot, there seems to be a strong positive correlation between petal length and sepal length. This means that as sepal length increases, it's likely for petal length to increase as well.


### Question 2

Based on the bar chart, the virginica species of iris has the greatest average petal length. Versicolor would be second and versicolor would be last.